% ==============================================================================
% Chapter 5 — Results: Ioannina
% ==============================================================================

\chapter{Application to Ioannina}
\label{ch:ioannina}

\todo[inline]{Write Ioannina results chapter.}

\section{Study Area}
\label{sec:ioannina_area}

\todo[inline]{Describe Ioannina: geography (northwest Greece, mountainous basin),
population ($\sim$65,000 inhabitants), main emission sources (residential
heating, road transport), known air quality issues (winter PM episodes from
wood burning).}

\section{Downscaled Emission Inventory}
\label{sec:ioannina_emissions}

\todo[inline]{Present the 1x1 km downscaled emission maps for Ioannina.
Include figures for NOx, PM2.5, PM10 by GNFR sector (road transport and
residential heating are the dominant sectors). Compare total emission budgets
between the original CAMS grid and the downscaled product.}

\subsection{Road Transport}
\label{sec:ioannina_road}

\todo[inline]{Show spatial distribution of road transport NOx emissions.
Discuss the impact of the updated road proxy (tertiary roads, OTM weights).}

\subsection{Residential Combustion}
\label{sec:ioannina_residential}

\todo[inline]{Discuss the allocation of residential heating emissions.
Note the importance of wood burning in the Ioannina context.}

\section{Air Quality Modelling Results}
\label{sec:ioannina_aqm}

\todo[inline]{Present EPISODE-CityChem concentration fields (NO2, PM2.5, PM10).
Show annual/seasonal mean maps. Compare modelled vs. observed concentrations
at monitoring stations.}

\subsection{Model Evaluation}
\label{sec:ioannina_eval}

\todo[inline]{Present statistical evaluation metrics (Table and scatter plots):
RMSE, bias, correlation, FAC2. Compare performance of updated vs. original
inventory.}

\section{Sensitivity to Inventory Version}
\label{sec:ioannina_sensitivity}

\todo[inline]{Compare model results using the original CAMS-REG v4 inventory vs.
the RI-URBANS updated inventory to quantify the impact of the improvements.}

\section{Climate Policy Co-Benefit Scenarios}
\label{sec:ioannina_scenarios}

\todo[inline]{Describe the mitigation scenarios defined in the Metronome project
for Ioannina (e.g., building energy efficiency measures, EV uptake in transport).
Present the resulting changes in emission totals and modelled air quality
concentrations.}
