% ==============================================================================
% Chapter 6 — Results: Kozani
% ==============================================================================

\chapter{Application to Kozani}
\label{ch:kozani}

\todo[inline]{Write Kozani results chapter.}

\section{Study Area}
\label{sec:kozani_area}

\todo[inline]{Describe Kozani: geography (northern Greece, Western Macedonia),
population ($\sim$40,000 inhabitants), dominant emission sources (heavy industry
/ lignite power plants in the broader region, road transport, residential
heating), known air quality challenges.}

\section{Downscaled Emission Inventory}
\label{sec:kozani_emissions}

\todo[inline]{Present the 1x1 km downscaled emission maps for Kozani.
Include comparison figures for NOx, PM2.5, PM10. Note the role of
industrial point sources (E-PRTR) in the Kozani domain vs. Ioannina.}

\subsection{Road Transport}
\label{sec:kozani_road}

\todo[inline]{Show spatial distribution of road transport NOx emissions.}

\subsection{Industrial Sources}
\label{sec:kozani_industry}

\todo[inline]{Discuss allocation of industrial (GNFR A and B) emissions using
E-PRTR point source data and CLC industrial areas.}

\subsection{Residential Combustion}
\label{sec:kozani_residential}

\todo[inline]{Discuss residential heating emissions.}

\section{Air Quality Modelling Results}
\label{sec:kozani_aqm}

\todo[inline]{Present EPISODE-CityChem concentration fields. Show annual/seasonal
mean maps. Compare modelled vs. observed concentrations.}

\subsection{Model Evaluation}
\label{sec:kozani_eval}

\todo[inline]{Present statistical evaluation metrics. Compare performance of
updated vs. original inventory.}

\section{Sensitivity to Inventory Version}
\label{sec:kozani_sensitivity}

\todo[inline]{Quantify the impact of the RI-URBANS inventory improvements on
modelled concentrations for Kozani.}

\section{Climate Policy Co-Benefit Scenarios}
\label{sec:kozani_scenarios}

\todo[inline]{Describe and present results of the Metronome mitigation scenarios
for Kozani. Discuss similarities and differences with Ioannina results.}
