% ==============================================================================
% Chapter 7 — Discussion and Conclusion
% ==============================================================================

\chapter{Discussion and Conclusion}
\label{ch:discussion}

\todo[inline]{Write discussion and conclusion chapter.}

\section{Discussion}
\label{sec:discussion}

\subsection{Impact of the Updated Emission Inventory}
\label{sec:disc_inventory}

\todo[inline]{Discuss the overall impact of the RI-URBANS inventory improvements
across both cities. Compare findings with those from the intercomparison study
in \citet{Kuenen_ST15_2025} (Helsinki, Amsterdam, Birmingham).}

\subsection{Spatial Proxy Performance}
\label{sec:disc_proxies}

\todo[inline]{Discuss the performance of the updated road transport proxy
(OTM weights, tertiary roads) and the new off-road sub-sector proxies.
Identify remaining uncertainties.}

\subsection{UFP Number Emissions}
\label{sec:disc_ufp}

\todo[inline]{Discuss the first application of TPN emission maps for Ioannina
and Kozani. Identify data gaps and recommended future improvements.}

\subsection{Comparison between the Two Cities}
\label{sec:disc_comparison}

\todo[inline]{Discuss the similarities and differences in emission structure
and model performance between Ioannina (dominated by residential combustion)
and Kozani (significant industrial sources).}

\subsection{Co-Benefit Analysis}
\label{sec:disc_cobenefit}

\todo[inline]{Synthesise the co-benefit results across both cities. Discuss
implications for local policy (building retrofit programmes, EV transition).}

\section{Limitations and Uncertainties}
\label{sec:limitations}

\todo[inline]{List main limitations: base year mismatch, uncertainty in activity
data, spatial proxy quality for small Greek cities, meteorological input
uncertainty, boundary condition uncertainty.}

\section{Conclusions}
\label{sec:conclusions}

\todo[inline]{Write conclusions (1--2 pages). Structured as numbered conclusions
corresponding to the objectives stated in Section~\ref{sec:intro_objectives}.}

\section{Outlook and Recommendations}
\label{sec:outlook}

\todo[inline]{Recommend: extension to additional Metronome cities, integration
of updated COPERT/OTM data, coupling with WRF meteorology, evaluation of
condensable PM fraction in residential emissions, further UFP inventory
development.}
